Holomorphic field theories and higher algebra

### 2022-12-01 draft.tex
tiny tweaks in a quick read

In studying questions about the classical vacua of supersymmetric field theories, however, physicists have not typically worked with the {\em stack}.
Instead, they have worked with a GIT quotient.
For instance, in the setting of a gauge theory with moduli stack of classical vacua $[V/G]$, it means they work with 
\[
V\Sslash G = \Spec(\cO(V)^G),
\]
the scheme whose ring of functions are the $G$-invariant functions on the representation~$V$.
This affine scheme is known as the {\em affinization} of the stack $V/G$, as it is the best affine approximation to the stack (i.e., apply the left adjoint to the inclusion of affine schemes among stacks).
It might be seen as parametrizing ``states'' rather than solutions, 
as it is defined via functions on the solutions (see Remark~\ref{sols vs states} for more).
We will call this GIT quotient the moduli {\em space} of classical vacua and denote it by~$M_{\rm vac}(\cT)$.
(Note that the font of M changes between stack and space.)

\begin{rmk}
In practice physicists work ordinary schemes (i.e., without any derived directions), 
although one can also study the derived affine scheme that is the best approximation to a derived stack.
Below we will always provide an ordinary scheme as the moduli {\em space}.
\end{rmk}

It is familiar to geometers that this affinization forgets interesting information about the stack because an invariant function must be constant on the closure of any orbit.
Thus, the canonical map
\[
[V/G] \to V\Sslash G
\]
often collapses orbits together.
An important example, pertinent to $\cN =1$ super Yang-Mills theory, is the adjoint quotient $[\gl_N/ GL_N]$.
Its affinization is
\[
\gl_N \Sslash GL_N \cong \CC^N
\]
by the Chevalley restriction theorem, 
which says $\cO(\gl_N)^{\GL_N} \cong \cO(\frak{h})^{S_N}$ where $\frak{h}$ denotes the maximal torus.
(The rightmost algebra is symmetric polynomials.)
The canonical map here records the characteristic polynomial of a matrix and hence only knows its generalized eigenvalues and not its Jordan decomposition.
In other words, the canonical map collapses together Jordan blocks with the same diagonal entries.
For this reason, it loses details about the classical vacua remembered by the stack quotient.

@@

For the holomorphic QCD theory with $F$ flavors and $N$ colors (without mesons), 
the moduli stack of classical vacua is 
\[
\cM_{\rm vac}(\cT_{\sfE}(F,N)) \cong T^*[1] \left(V_Q \times V_Q^*/SL(N)\right).
\]
By contrast the moduli space $M_{\rm vac}$ is a variety already well-studied in classical algebraic geometry.
Before we describe the full space, let's examine a simpler case involving just the quarks.
A vacuum for the quark field is an element of $V_Q = \Hom(\CC^F, \CC^N)$,
and it is a classic result of invariant theory that
\[
\Hom(\CC^F, \CC^N)\Sslash SL(N) \cong \widehat{{\rm Gr}(F, N)} \subset \Lambda^N \CC^F
\]
where ${\rm Gr}(F, N)$ denotes the Grassmannian of $N$-planes in $\CC^F$ and the hat indicates we take the affine cone with respect to this standard Pl\"ucker embedding.
For the full problem, we have both quarks and antiquarks, so
\begin{align*}
M_{\rm vac}(\cT_{\sfE}(F,N)) 
&\cong V_Q \times V_Q^*\Sslash SL(N) \\
&= \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)\Sslash SL(N),
\end{align*}
whose ring of functions is also computed by the fundamental theorems of invariant theory.
Those theorems tell us that the GIT quotient is a closed conical subvariety
\[
M_{\rm vac}(\cT_{\sfE}(F,N))  \hookrightarrow \Lambda^N \CC^F \times \Lambda^N (\CC^F)^* \times \End(\CC^F)
\]
where a pair of matrices $(X,Y) \in \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)$ goes to $(\Lambda^N X, \Lambda^N Y, YX)$, with $\Lambda^N X$ denoting the wedge product of the rows of $X$ (and of columns for $Y$).
The image thus satisfies both sets of Pl\"ucker relations, the condition that the composite $YX$ has rank at most $N$, and some further relations, given by the Second Fundamental Theorem, such as that the pairing of $\Lambda^N X$ and $\Lambda^N Y$ can be related to $\Lambda^N (YX)$.
(See Chapter 2, Section C of \cite{Weyl}.)
In other words, we have a closed subvariety
\[
M_{\rm vac}(\cT_{\sfE}(F,N)) \hookrightarrow \widehat{{\rm Gr}(F, N)} \times \widehat{{\rm Gr}(F, N)} \times {\rm R}(F,N)
\]
with ${\rm R}(F,N) \subset \End(\CC^F)$ the matrices with rank bounded above by~$N$.
The moduli space for the magnetic theory is a bit more complicated but follows from classical invariant theory as well.

Note that $\check{N} = F-N$, so $\widehat{{\rm Gr}(F, N)} \cong \widehat{{\rm Gr}(F, \check{N})}$,
but ${\rm R}(F,N)$ is not isomorphic to ${\rm R}(F,\check{N})$ except when $N = \check{N}$.
Hence one should not be optimistic that $M_{\rm vac}(\cT_{\sfE}(F,N)) \cong M_{\rm vac}(\cT_{\sfE}(F,N))$.
To ``fix'' this issue, one might try to slightly modify the dual side, which Seiberg did by adding the meson fields.
Hence the following holds.

@@

\begin{rmk}
In the literature this isomorphism is given by explicitly matching gauge-invariant functions, 
which are often called mesons or baryons, 
as they can be identified with the data that identifies such particles.
These meson and baryons are precisely the usual generators of the ring of invariant functions, 
so an isomorphism of generators determines the desired isomorphism of rings. 
One must check that the relations are satisfied;
this requirement explains the superpotential on the magnetic side.
\end{rmk}

@@

\subsubsection{$SU(2)$ SQCD}

>>>>

In studying questions about the classical vacua of supersymmetric field theories, however, physicists have not typically worked with the {\em stack}.
Instead, they have worked with a GIT quotient.
For instance, in the setting of a gauge theory with moduli stack of classical vacua $[V/G]$, it means they work with 
\[
V\Sslash G = \Spec(\cO(V)^G),
\]
the scheme whose ring of functions are the $G$-invariant functions on the representation~$V$.
This affine scheme is known as the {\em affinization} of the stack $V/G$, as it is the best affine approximation to the stack (i.e., apply the left adjoint to the inclusion of affine schemes among stacks).
This affinization might be seen as parametrizing ``states'' rather than solutions, 
as it is defined via functions on the solutions (see Remark~\ref{sols vs states} for more on this aspect).

We will call this GIT quotient the moduli {\em space} of classical vacua and denote it by~$M_{\rm vac}(\cT)$.
Note that the font of M changes between stack and space.

\begin{rmk}
In practice physicists work with ordinary schemes (i.e., without any derived directions), 
although one can also study the derived affine scheme that is the best approximation to a derived stack.
Below we will always provide an ordinary scheme as the moduli {\em space},
following the physicists.
\end{rmk}

It is familiar to geometers that this affinization forgets interesting information about the stack because an invariant function must be constant on the closure of any orbit.
The canonical map
\[
[V/G] \to V\Sslash G
\]
often collapses orbits together.
An important example, pertinent to $\cN =1$ super Yang-Mills theory, is the adjoint quotient $[\gl_N/ GL_N]$.
Its affinization is
\[
\gl_N \Sslash GL_N \cong \CC^N
\]
by the Chevalley restriction theorem, 
which says $\cO(\gl_N)^{\GL_N} \cong \cO(\frak{h})^{S_N}$ where $\frak{h}$ denotes the maximal torus.
(The rightmost algebra is symmetric polynomials in the eigenvalues.)
The canonical map here records the characteristic polynomial of a matrix and hence only knows its generalized eigenvalues and not its Jordan decomposition.
In other words, the canonical map collapses together Jordan blocks with the same diagonal entries.
For this reason, it loses details about the classical vacua remembered by the stack quotient.

@@

For the holomorphic QCD theory with $F$ flavors and $N$ colors (without mesons), 
the moduli stack of classical vacua is 
\[
\cM_{\rm vac}(\cT_{\sfE}(F,N)) \cong T^*[1] \left(V_Q \times V_Q^*/SL(N)\right).
\]
By contrast the moduli space $M_{\rm vac}$ is a variety already well-studied in classical algebraic geometry.
Before we describe the full space, let's examine a simpler case involving just the quarks.
A vacuum for the quark field is an element of $V_Q = \Hom(\CC^F, \CC^N)$,
and it is a classic result of invariant theory that
\[
\Hom(\CC^F, \CC^N)\Sslash SL(N) \cong \widehat{{\rm Gr}(F, N)} \subset \Lambda^N \CC^F
\]
where ${\rm Gr}(F, N)$ denotes the Grassmannian of $N$-planes in $\CC^F$ and the hat indicates we take the affine cone with respect to this standard Pl\"ucker embedding.

For the full problem, we have both quarks and antiquarks, so
\begin{align*}
M_{\rm vac}(\cT_{\sfE}(F,N)) 
&\cong V_Q \times V_Q^*\Sslash SL(N) \\
&= \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)\Sslash SL(N),
\end{align*}
whose ring of functions is also computed by the fundamental theorems of invariant theory.
Those theorems tell us that the GIT quotient is a closed conical subvariety
\[
M_{\rm vac}(\cT_{\sfE}(F,N))  \hookrightarrow \Lambda^N \CC^F \times \Lambda^N (\CC^F)^* \times \End(\CC^F)
\]
where a pair of matrices $(X,Y) \in \Hom(\CC^F, \CC^N) \times \Hom(\CC^N, \CC^F)$ goes to $(\Lambda^N X, \Lambda^N Y, YX)$, with $\Lambda^N X$ denoting the wedge product of the rows of $X$ (and of columns for $Y$).
The image thus satisfies both sets of Pl\"ucker relations, the condition that the composite $YX$ has rank at most $N$, and some further relations, given by the Second Fundamental Theorem, such as that the pairing of $\Lambda^N X$ and $\Lambda^N Y$ can be related to $\Lambda^N (YX)$.
(See Chapter 2, Section C of \cite{Weyl}.)
In other words, we have a closed subvariety
\[
M_{\rm vac}(\cT_{\sfE}(F,N)) \hookrightarrow \widehat{{\rm Gr}(F, N)} \times \widehat{{\rm Gr}(F, N)} \times {\rm R}(F,N)
\]
with ${\rm R}(F,N) \subset \End(\CC^F)$ the matrices with rank bounded above by~$N$.
The moduli space for the magnetic theory is a bit more complicated but follows from classical invariant theory as well.

Note that $\check{N} = F-N$, so ${\rm Gr}(F, N) \cong {\rm Gr}(F, \check{N})$ and a (partial) duality is apparent.
On the other hand, ${\rm R}(F,N)$ is not isomorphic to ${\rm R}(F,\check{N})$ except when $N = \check{N}$.
Hence one should not be optimistic that $M_{\rm vac}(\cT_{\sfE}(F,N)) \cong M_{\rm vac}(\cT_{\sfE}(F,N))$.
To ``fix'' this issue, one might try to slightly modify the dual side, which Seiberg did by adding the meson fields.
Hence the following holds.

@@

\begin{rmk}
In the literature this isomorphism is given by explicitly matching gauge-invariant functions, 
which physics paper often call mesons or baryons, 
as they can be identified with the data that identifies such particles.
These meson and baryons are precisely the usual generators of the ring of invariant functions, 
so an isomorphism of generators determines the desired isomorphism of rings. 
One must check that the relations are satisfied;
this requirement explains the superpotential on the magnetic side.
\end{rmk}

@@

\subsubsection{The special case of $SU(2)$ SQCD}
